---
title: "現代の武力紛争"
subtitle: "政治学入門"
author: "土井翔平"
date: today
---

```{r}
#| include: false
#| cache: false

library(tidyverse)
theme_set(theme_bw(base_family = "Noto Sans CJK JP Light"))
```

## 導入 {.unnumbered}

- 現代の武力紛争の特徴はなにか？

武力紛争を定義し、[データセット](https://www.cao.go.jp/pko/pko_j/organization/researcher/atpkonow/article099.html)を構築

:::{.callout-note}
## [UCDPにおける武力紛争の定義](https://ucdp.uu.se/downloads/ucdpprio/ucdp-prio-acd-221.pdf)

UCDP defines state-based armed conflict as: “a **contested incompatibility** that concerns government and/or territory where the **use of armed force** between **two parties**, of which at least one is the government of a state, results in at least **25 battle-related deaths** in a calendar year.”
:::

::: {.column-screen-right}
<iframe src="https://ucdp.uu.se/encyclopedia" loading="lazy" style="width: 100%; height: 800px; border: 0px none;"></iframe>
:::

### 課題文献 {.unnumbered}

- 戦争のデータ
  - @tago2020[序章]

## 武力紛争の頻度

```{r}
#| fig-cap: "武力紛争 (UCDP) の発生件数"

df_ucdp <- read_csv("data/ucdp-prio-acd-221.csv")

df_ucdp %>% 
  count(year) %>% 
  ggplot() + 
  geom_line(aes(x = year, y = n), size = 0.25) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  labs(x = "年", y = "武力紛争発生数")
```

```{r}
#| fig-cap: "武力紛争 (UCDP) の種類"

type <- c("独立戦争", "国家間戦争", "内戦", "国際介入のある内戦")

left_join(expand(df_ucdp, year, type_of_conflict),
          df_ucdp %>% 
            count(year, type_of_conflict)) %>% 
  replace_na(list(n = 0)) %>% 
  group_by(year) %>% 
  mutate(total = sum(n),
         ratio = n/total) %>% 
  pivot_longer(c(n, ratio)) %>% 
  mutate(name = recode(name, n = "件数", ratio = "比率")) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = value, 
                color = as.factor(type_of_conflict),
                fill = as.factor(type_of_conflict)),
            alpha = 0.5) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  facet_grid(name ~ ., scales = "free_y") + 
  labs(x = "年", y = "武力紛争", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

[Correlates of War (COW)](https://correlatesofwar.org/data-sets/MIDs) のデータ[@palmer2022]

- No militarized action
- Threat to use force
- Display use of force
- Use of force, War

```{r}
#| fig-cap: "国家間武力衝突 (MID) の発生件数"

df_mid_temp <- read_csv("data/mid/MIDA 5.0.csv", na = "-9")

df_mid <- NULL

for (i in 1:nrow(df_mid_temp)) {
  temp <- df_mid_temp[i,]
  year <- temp$styear:temp$endyear
  df_mid <- bind_rows(df_mid, tibble(dispnum = temp$dispnum, year = year))
}

df_mid <- left_join(df_mid, df_mid_temp) %>% 
  mutate(intensity_level = case_when(fatality == 0 ~ "0", 
                                     fatality == 1 ~ "1",
                                     between(fatality, 2, 5) ~ "2", 
                                     fatality == 6 ~ "3",
                                     TRUE ~ "4"))

df_mid %>% 
  count(year) %>% 
  ggplot() + 
  geom_line(aes(x = year, y = n), size = 0.25) + 
  annotate("rect", xmin = 1914, xmax = 1918, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  annotate("rect", xmin = 1939, xmax = 1945, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  labs(x = "年", y = "武力紛争発生数")
```

```{r}
#| fig-cap: "国家間武力衝突 (MID) の種類"

type <- c("武力による威嚇", "武力の動員", "武力行使", "戦争")

left_join(expand(df_mid, year, hostlev),
          df_mid %>% 
            count(year, hostlev)) %>% 
  replace_na(list(n = 0)) %>% 
  group_by(year) %>% 
  mutate(total = sum(n),
         ratio = n/total) %>% 
  pivot_longer(c(n, ratio)) %>% 
  mutate(name = recode(name, n = "件数", ratio = "比率")) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = value, 
                color = as.factor(hostlev),
                fill = as.factor(hostlev)),
            alpha = 0.5, size = 0.25) + 
  annotate("rect", xmin = 1914, xmax = 1918, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  annotate("rect", xmin = 1939, xmax = 1945, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  facet_grid(name ~ ., scales = "free_y") + 
  labs(x = "年", y = "武力紛争発生数", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "国家間武力衝突 (MID) の当事国の割合"

df_system <- read_csv("data/system2016.csv")

df_midb_temp <- read_csv("data/mid/MIDB 5.0.csv", na = "-9")

df_midb <- NULL

for (i in 1:nrow(df_midb_temp)) {
  temp <- df_midb_temp[i,]
  year <- temp$styear:temp$endyear
  df_midb <- bind_rows(df_midb, tibble(dispnum = temp$dispnum, 
                                       stateabb = temp$stabb, 
                                       year = year))
}

left_join(df_system,
          df_midb %>% 
            distinct(stateabb, year) %>% 
            mutate(mid = 1)) %>% 
  filter(year < 2015) %>% 
  group_by(year) %>% 
  summarise(n_country = n(),
            n_dispute = sum(mid, na.rm = TRUE)) %>% 
  mutate(ratio = n_dispute/n_country) %>% 
  ggplot() + 
  geom_line(aes(x = year, y = ratio), size = 0.2) + 
  annotate("rect", xmin = 1914, xmax = 1918, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  annotate("rect", xmin = 1939, xmax = 1945, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  geom_hline(aes(yintercept = mean(ratio, na.rm = TRUE)), 
             linetype = "dashed", alpha = 0.5) + 
  labs(x = "年", y = "MIDの当事国の比率")
```

## 武力紛争の空間

```{r}
#| fig-cap: "武力紛争の分布（1648-2003年）[@miller2023]"

df_hced <- read_csv("data/HCED Data v2.csv") %>% 
  drop_na(ID) %>% 
  mutate(Year = parse_number(Year))

df_world <- map_data("world") %>% 
  filter(region != "Antarctica")

df_world %>% 
  ggplot() + 
  geom_polygon(aes(x = long, y = lat, group = group), 
               fill = "lightgray", color = "black", linewidth = 0.05, alpha = 0.25) + 
  geom_point(aes(x = Longitude, y = Latitude, color = Year), data = df_hced %>% filter(Year >= 1648), 
             size = 0.05, alpha = 0.5) +
  scale_color_viridis_c() + 
  coord_map(projection = "mollweide", xlim = c(-180, 180)) + 
  labs(x = "", y = "", color = "") + 
  theme_void()
```

## 武力紛争の理由

```{r}
#| fig-cap: "武力紛争 (UCDP) の対立理由"

type <- c("領土", "政府", "領土及び政府")

left_join(expand(df_ucdp, year, incompatibility),
          df_ucdp %>% 
            count(year, incompatibility)) %>% 
  replace_na(list(n = 0)) %>% 
  group_by(year) %>% 
  mutate(total = sum(n),
         ratio = n/total) %>% 
  pivot_longer(c(n, ratio)) %>% 
  mutate(name = recode(name, n = "件数", ratio = "比率")) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = value, 
                color = as.factor(incompatibility),
                fill = as.factor(incompatibility)),
            alpha = 0.5) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  facet_grid(name ~ ., scales = "free_y") + 
  labs(x = "年", y = "武力紛争", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

## 武力紛争の被害

```{r}
#| fig-cap: "武力衝突 (UCDP) の規模"

type <-  c("25人以上1000人未満", "1000人以上")

left_join(expand(df_ucdp, year, intensity_level),
          df_ucdp %>% 
            count(year, intensity_level)) %>% 
  replace_na(list(n = 0)) %>% 
  group_by(year) %>% 
  mutate(total = sum(n),
         ratio = n/total) %>% 
  pivot_longer(c(n, ratio)) %>% 
  mutate(name = recode(name, n = "件数", ratio = "比率")) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = value, 
                color = as.factor(intensity_level),
                fill = as.factor(intensity_level)),
            alpha = 0.5) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  facet_grid(name ~ ., scales = "free_y") + 
  labs(x = "年", y = "武力紛争", 
       color = "武力紛争に関する死者数", fill = "武力紛争に関する死者数") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "国家間軍事衝突 (COW) の規模"

type <- c("0人", "0人以上25人未満", "25人以上1000人未満", "1000人以上", "不明")

left_join(expand(df_mid, year, intensity_level),
          df_mid %>% 
            count(year, intensity_level)) %>% 
  replace_na(list(n = 0)) %>% 
  replace_na(list(n = 0)) %>% 
  group_by(year) %>% 
  mutate(total = sum(n),
         ratio = n/total) %>% 
  pivot_longer(c(n, ratio)) %>% 
  mutate(name = recode(name, n = "件数", ratio = "比率")) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = value, 
                color = as.factor(intensity_level),
                fill = as.factor(intensity_level)),
            alpha = 0.5, size = 0.25) + 
  annotate("rect", xmin = 1914, xmax = 1918, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  annotate("rect", xmin = 1939, xmax = 1945, ymin = -Inf, ymax = Inf, alpha = 0.25) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  facet_grid(name ~ ., scales = "free_y") + 
  labs(x = "年", y = "武力紛争発生数", 
       color = "武力紛争に関する死者数", fill = "武力紛争に関する死者数") + 
  theme(legend.position = "bottom") + 
  guides(color = guide_legend(nrow = 2), fill = guide_legend(nrow = 2))
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と被害者数の推移"

df_brd <- read_csv("data/ucdp-brd-dyadic-221.csv")

type <- c("国家間戦争", "内戦", "国際介入のある内戦")

left_join(df_brd %>% 
            filter(year > 1988) %>% 
            expand(year, type_of_conflict),
          df_brd %>% 
            group_by(year, type_of_conflict) %>% 
            summarise(bd_total = sum(bd_best, na.rm = TRUE))) %>% 
  replace_na(list(bd_total = 0)) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = bd_total, 
                color = as.factor(type_of_conflict), 
                fill = as.factor(type_of_conflict)), 
            alpha = 0.5) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "年", y = "武力紛争に関する死者数", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と被害者数"

df_brd %>% 
  filter(year > 1988) %>% 
  group_by(conflict_id, type_of_conflict) %>% 
  summarise(bd_total = sum(bd_best, na.rm = TRUE)) %>% 
  ggplot() + 
  geom_histogram(aes(x = bd_total, 
                     color = as.factor(type_of_conflict),
                     fill = as.factor(type_of_conflict)), 
                 alpha = 0.5) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争に関する死者数", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と被害者数"

df_brd %>% 
  filter(year > 1988) %>% 
  group_by(conflict_id, type_of_conflict) %>% 
  summarise(bd_total = sum(bd_best, na.rm = TRUE)) %>% 
  ggplot() + 
  geom_histogram(aes(x = bd_total, 
                     color = as.factor(type_of_conflict),
                     fill = as.factor(type_of_conflict)), 
                 alpha = 0.5, position = position_fill()) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争に関する死者数", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の対立理由と被害者数の推移"

type <-  c("領土", "政府", "領土及び政府")

left_join(df_brd %>% 
            filter(year > 1988) %>% 
            expand(year, incompatibility),
          df_brd %>% 
            group_by(year, incompatibility) %>% 
            summarise(bd_total = sum(bd_best, na.rm = TRUE))) %>% 
  replace_na(list(bd_total = 0)) %>% 
  ggplot() + 
  geom_area(aes(x = year, y = bd_total, 
                color = as.factor(incompatibility), 
                fill = as.factor(incompatibility)), 
            alpha = 0.5) + 
  geom_vline(aes(xintercept = 1991), alpha = 0.25) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "年", y = "武力紛争に関する死者数", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と被害者数"

df_brd %>% 
  filter(year > 1988) %>% 
  group_by(conflict_id, incompatibility) %>% 
  summarise(bd_total = sum(bd_best, na.rm = TRUE)) %>% 
  ggplot() + 
  geom_histogram(aes(x = bd_total, 
                     color = as.factor(incompatibility),
                     fill = as.factor(incompatibility)), 
                 alpha = 0.5) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争に関する死者数", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と被害者数"

df_brd %>% 
  filter(year > 1988) %>% 
  group_by(conflict_id, incompatibility) %>% 
  summarise(bd_total = sum(bd_best, na.rm = TRUE)) %>% 
  ggplot() + 
  geom_histogram(aes(x = bd_total, 
                     color = as.factor(incompatibility),
                     fill = as.factor(incompatibility)), 
                 alpha = 0.5, position = position_fill()) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争に関する死者数", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

## 武力紛争の期間

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と期間"

type <- c("独立戦争", "国家間戦争", "内戦", "国際介入のある内戦")

df_ucdp %>% 
  distinct(conflict_id, start_date2, ep_end_date, .keep_all = TRUE) %>% 
  replace_na(list(ep_end_date = as.Date("2020-12-31"))) %>% 
  mutate(length = time_length(interval(start_date2, ep_end_date), "day") + 1,
         year = year(start_date2)) %>% 
  ggplot() + 
  geom_histogram(aes(x = length/365, 
                     color = as.factor(type_of_conflict),
                     fill = as.factor(type_of_conflict)), 
                 alpha = 0.5) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争の期間（年）", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の形態と期間"

type <- c("独立戦争", "国家間戦争", "内戦", "国際介入のある内戦")

df_ucdp %>% 
  distinct(conflict_id, start_date2, ep_end_date, .keep_all = TRUE) %>% 
  replace_na(list(ep_end_date = as.Date("2020-12-31"))) %>% 
  mutate(length = time_length(interval(start_date2, ep_end_date), "day") + 1,
         year = year(start_date2)) %>% 
  ggplot() + 
  geom_histogram(aes(x = length/365, 
                     color = as.factor(type_of_conflict),
                     fill = as.factor(type_of_conflict)), 
                 alpha = 0.5, position = position_fill()) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争の期間（年）", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の対立原因と期間"

type <-  c("領土", "政府", "領土及び政府")

df_ucdp %>% 
  distinct(conflict_id, start_date2, ep_end_date, .keep_all = TRUE) %>% 
  replace_na(list(ep_end_date = as.Date("2020-12-31"))) %>% 
  mutate(length = time_length(interval(start_date2, ep_end_date), "day") + 1,
         year = year(start_date2)) %>% 
  ggplot() + 
  geom_histogram(aes(x = length/365, 
                     color = as.factor(incompatibility),
                     fill = as.factor(incompatibility)), 
                 alpha = 0.5) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争の期間（年）", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```

```{r}
#| fig-cap: "武力衝突 (UCDP) の対立原因と期間"

type <-  c("領土", "政府", "領土及び政府")

df_ucdp %>% 
  distinct(conflict_id, start_date2, ep_end_date, .keep_all = TRUE) %>% 
  replace_na(list(ep_end_date = as.Date("2020-12-31"))) %>% 
  mutate(length = time_length(interval(start_date2, ep_end_date), "day") + 1,
         year = year(start_date2)) %>% 
  ggplot() + 
  geom_histogram(aes(x = length/365, 
                     color = as.factor(incompatibility),
                     fill = as.factor(incompatibility)), 
                 alpha = 0.5, position = position_fill()) + 
  scale_color_brewer(palette = "Set2", labels = type) + 
  scale_fill_brewer(palette = "Set2", labels = type) + 
  labs(x = "武力紛争の期間（年）", y = "", color = "", fill = "") + 
  theme(legend.position = "bottom")
```
